### 2017-10-26 inst/miR34aLocus/miR34aLocus.Rmd

---
title: "miR34a locus"
author: "Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    toc: no
    code_folding: hide
---

#Introduction
An antisense transcript to the miR34a HG was detected using the UCSC genome browser.

#Methods
The USCS genome browser was utilized for the bioinformatic evaluation of antisense transcription utilizing the Human EST and ENCODE RNAseq tracks.

#Results
<br></br>
<center>![miR34a locus and cloning of P1 and P2 constructs.](miR34aLocus_hg38.png)</center>
<br></br>
A schematic picture of the miR34a locus from UCSC genome browser including miR34a HG and mature miR34a and LINC01759. H3K4me3 ChIP-seq data from ENCODE indicating the promoter region and conservation over the locus is also shown.

#Conclusion
GENCODE annotation supports the presence of a transcript orinionating in the antisense orientation from the miR34a gene.