### 2017-11-07 inst/miR34aLocus/figure1a.Rmd

---
title: "miR34a locus"
author: "Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    toc: no
    code_folding: hide
---

####Introduction
An antisense transcript to the miR34a HG was detected using the UCSC genome browser.

####Methods
The USCS genome browser was utilized for the bioinformatic evaluation of antisense transcription utilizing the Human EST and ENCODE RNAseq tracks.

####Results
```{r, out.width = "800px", tidy=TRUE, tidy.opts=list(width.cutoff=60)}
projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst/"
url <- "miR34aLocus/miR34aLocus_hg38.png"
knitr::include_graphics(file.path(projectUrl, url))
```
A schematic picture of the miR34a locus from UCSC genome browser (hg38) including miR34a HG and mature miR34a, and LINC01759. H3K4me3 ChIP-seq data and conservation over the locus is also shown.

####Conclusion
GENCODE annotation supports the presence of a transcript orinionating in the antisense orientation from the miR34a gene.