### 2017-11-12 inst/lnc34aSplJnc/lnc34aSplJnc.Rmd
more accurate counting of splJnc reads in analysis

---
title: "lnc34a splice junctions"
author: "Jason T. Serviss"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
packages <- c(
    "tidyverse",
    "grid",
    "gtable",
    "miR34AasRNAproject"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)
```

```{r}
#locus
data <- tibble(
  gene = readr::parse_factor(
    c(rep("miR34a asRNA", 5), rep("lnc34a", 3)),
    levels = c("miR34a asRNA", "lnc34a")
  ),
  feature = c(
    "exon", "intron", "exon", "intron",
    "exon", "exon", "intron", "exon"
  ),
  start = c(
    9242262, 9242375, 9243691, 9243866,
    9251451, 9241796, 9242354, 9256971),
  stop = c(
    9242375, 9243691, 9243866, 9251451,
    9252148, 9242354, 9256971, 9257102
  )
)

#splice junctions
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/data"
url <- "lnc34aSpliceJncs.rds"
splJnc <- read_rds(gzcon(url(file.path(dataUrl, url))))

splJnc <- splJnc %>%
  mutate(cellline = gsub(".*LongRnaSeq(.*)Cell.*", "\\1", filename)) %>%
  select(cellline, start, stop, score2) %>%
  group_by(start, stop) %>%
  summarize(n = sum(score2)) %>%
  ungroup() %>%
  filter(n > 2) %>%
  arrange(desc(n)) %>%
  mutate(y = 1:length(.data$n))
```

```{r, fig.align='center', fig.height=8, fig.width=12}
###locus
p1 <- ggplot(data = NULL) +
  gggenes::geom_gene_arrow(
    data = filter(data, feature == "exon"),
    aes(xmin = start, xmax = stop, y = gene, fill = gene),
    arrowhead_width = unit(0, "mm"),
    arrowhead_height = unit(0, "mm"),
    colour = "gray"
  ) +
  ggthemes::theme_few() +
  ggthemes::scale_fill_ptol() +
  geom_segment(
    data = filter(data, feature == "intron"),
    aes(x = start, xend = stop, y = gene, yend = gene),
    linetype = "dotted",
    colour = "gray",
    size = 1
  ) +
  xlim(9241000, 9257900) +
  labs(x = "Chromosome 1", y = "Genes") +
  guides(fill = FALSE)

##splice junctions
p2 <- ggplot(data = NULL) +
  gggenes::geom_gene_arrow(
    data = splJnc,
    aes(xmin = start, xmax = stop, y = y, fill = n),
    arrowhead_width = unit(0, "mm"),
    arrowhead_height = unit(0, "mm"),
    arrow_body_height = unit(1, "mm"),
    colour = "gray"
  ) +
  viridis::scale_fill_viridis() +
  ggthemes::theme_few() +
  theme(
    axis.title.x = element_blank(),
    legend.position = "top",
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(y = "Splice junctions") +
  xlim(9241000, 9257900) +
  guides(fill = guide_colourbar(title = "n reads", title.position = "top", title.hjust = 0.5))

g2 <- ggplotGrob(p1)
g3 <- ggplotGrob(p2)
g <- rbind(g3, g2, size = "last")
grid.newpage()
grid.draw(g)

```