### 2017-11-12 inst/lnc34aCage/lnc34aCAGE.Rmd
lnc34a figures done in inst

---
title: "lnc34a CAGE"
author: "Jason T. Serviss"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Add number of cell lines included in figure.

```{r, message = FALSE}
packages <- c(
    "tidyverse",
    "grid",
    "gtable",
    "miR34AasRNAproject"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)
```

```{r}
data <- tibble(
  gene = readr::parse_factor(
    c(rep("miR34a asRNA", 5), rep("lnc34a", 3)),
    levels = c("miR34a asRNA", "lnc34a")
  ),
  feature = c(
    "exon", "intron", "exon", "intron",
    "exon", "exon", "intron", "exon"
  ),
  start = c(
    9242262, 9242375, 9243691, 9243866,
    9251451, 9241796, 9242354, 9256971),
  stop = c(
    9242375, 9243691, 9243866, 9251451,
    9252148, 9242354, 9256971, 9257102
  )
)

dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/data"
url <- "lnc34aCAGE.rds"
cage <- read_rds(gzcon(url(file.path(dataUrl, url)))) %>%
    arrange(desc(level)) %>%
    mutate(y = 1:length(pull(., level)))

```

```{r, fig.align='center', fig.height=8, fig.width=12}
###plot

p1 <- ggplot(data = NULL) +
  gggenes::geom_gene_arrow(
    data = filter(data, feature == "exon"),
    aes(xmin = start, xmax = stop, y = gene, fill = gene),
    arrowhead_width = unit(0, "mm"),
    arrowhead_height = unit(0, "mm"),
    colour = "gray"
  ) +
  ggthemes::theme_few() +
  ggthemes::scale_fill_ptol() +
  geom_segment(
    data = filter(data, feature == "intron"),
    aes(x = start, xend = stop, y = gene, yend = gene),
    linetype = "dotted",
    colour = "gray",
    size = 1
  ) +
  xlim(9241000, 9257900) +
  labs(x = "Chromosome 1", y = "Genes") +
  guides(fill = FALSE)

p2 <- ggplot(data = NULL) +
  gggenes::geom_gene_arrow(
    data = cage,
    aes(xmin = start, xmax = stop, y = y, fill = level),
    arrowhead_width = unit(0, "mm"),
    arrowhead_height = unit(0, "mm"),
    arrow_body_height = unit(2, "mm"),
    colour = "gray",
    show.legend = TRUE
  ) +
  ggthemes::theme_few() +
  viridis::scale_fill_viridis() +
  theme(
    axis.title.x = element_blank(),
    legend.position = "top",
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(y = "Cage reads") +
  xlim(9241000, 9257900) +
  guides(fill = guide_colourbar(title = "RPKM", title.position = "top", title.hjust = 0.5))

p3 <- ggplot(data = NULL) + 
  geom_density(data = cage, aes(x = start)) +
  ggthemes::theme_few() +
  xlim(9241000, 9257900) +
  theme(
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(y = "Density")

grobz <- lapply(list(p1, p2, p3), ggplotGrob)
g <- rbind(grobz[[2]], grobz[[3]], grobz[[1]], size = "last")
grid.newpage()
grid.draw(resize_heights(g, unit(c(3,1,1), "null")))
```