### 2017-11-14 inst/figure3b_stableLineCellCycle/figure3b_stableLineCellCycle.Rmd

---
title: "Figure 3b"
subtitle: "miR34a asRNA stable cell line cell cycle"
author: "Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}

packages <- c(
    "tidyverse",
    "printr",
    "ggthemes",
    "readr",
    "miR34AasRNAproject",
    "grid",
    "gtable"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst/"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/data"

```

####_Introduction_

miR34a is known to regulate cell cycle and, specifically, induce G1-phase arrest. Due to the increase observed in miR34a expression levels in cell lines over-expressing miR34a asRNA, we investigated if the cell cycle was perturbed in these cells.

<br></br>

####_Methods_

<br></br>

_Cell culture and cell cycle analysis_

All cell lines were cultured at 5% CO2 and 37° C with Skov3 and PC3 cells in RPMI (Hyclone) and 2 mM L-glutamine. All growth mediums were supplemented with 10% heat-inactivated FBS and 50 μg/ml of streptomycin and 50 μg/ml of penicillin. 1x10^5 cells per well in a 6-well plate and harvested after 24hrs. Cells were washed in PBS and fixed in 4% PFA at room temperature overnight. PFA was removed, and cells were resuspended in 95% EtOH. The samples were then rehydrated in distilled water, stained with DAPI and analyzed by flow cytometry on a LSRII (BD Biosciences) machine.

<br></br>

_Analysis_
Write the software used to quantify the cell cycle phases.

Percentages were converted to log2 fractions for each sample. Student's t-test was used for statistical analysis comparing the mock vs miR34a asRNA overespression conditions using five (PC3) or three (Skov3) independant experiments.

####_Results_

```{r, message = FALSE}

url <- "figure3b_stableLineCellCycle.rds"
data <- read_rds(gzcon(url(file.path(dataUrl, url))))

#convert the value (precent) to log2(fraction)
data <- mutate(data, value = log2(value / 100))

#qqplot shows that the data is reasonably normal. The shapiro test is also used below. Note that a p < 0.05 indicates non-normality.
#qqnorm(data$value)

#calculate stats
stats <- data %>%
  full_join(
    filter(data, condition == "mock"), 
    by = c("Biological Replicate", "Cell line", "phase"), 
    suffix = c("", ".y")
  ) %>%
  group_by(`Cell line`, condition, phase) %>%
  summarize(
    n = n(),
    mean = (2^mean(.data$value)),
    max = (2^max(.data$value)),
    CI95l = (2^t.test(.data$value)$conf.int[1]),
    CI95h = (2^t.test(.data$value)$conf.int[2]),
    pValue = t.test(.data$value, .data$value.y)$p.value,
    shapiroTest = shapiro.test(value)$p.value
  ) %>%
  ungroup() %>%
  pFormat(.)

#prepare for plotting
#change value back to percent
data <- data %>%
  mutate(value = 2^value)

#calculate start and end for horizontal plot lines
stats <- stats %>%
  group_by(phase, `Cell line`) %>%
  mutate(y = max(c(.data$CI95h, .data$max))) %>%
  ungroup() %>%
  arrange(`Cell line`, phase) %>%
  mutate(
    x = rep(sort(rep(1:3, 2)), 2) - 0.25,
    xend = rep(sort(rep(1:3, 2)), 2) + 0.25
  )

```

```{r, fig.align='center', fig.height=8, fig.width=12}

p <- ggplot(data = NULL) +
  geom_violin(
    data = data,
    aes(
      x = phase, 
      y = value,
      group = interaction(phase, condition)
    ),
    scale = "width",
    alpha = 0.3,
    fill = "gray",
    weight = 10,
    trim = TRUE,
    color = NA,
    show.legend = FALSE
  ) +
  geom_dotplot(
    data = data, 
    aes(
      x = phase,
      y = value,
      fill = condition
    ),
    binwidth = 1/100,
    colour = NA,
    alpha = 0.5,
    position = position_dodge(width = 0.9), 
    binaxis = 'y', 
    stackdir = 'center',
    dotsize = 0.5,
    show.legend = TRUE
  ) +
  facet_grid(
    . ~ `Cell line`, 
    scale = "free"
  ) +
  geom_linerange(
    data = stats,
    aes(
      x = phase, 
      ymin = CI95l, 
      ymax = CI95h, 
      group = interaction(phase, condition)
    ),
    position = position_dodge(width = 0.9),
    colour = "grey25",
    show.legend = FALSE
  ) +
  geom_point(
    data = stats,
    aes(
      x = phase, 
      y = mean, 
      group = condition
    ), 
    colour = "black",
    position = position_dodge(width = 0.9),
    shape = 95,
    size = 6,
    show.legend = FALSE
  ) +
  geom_segment(
    data = stats,
    aes(
      x = x,
      y = y + 0.01, 
      xend = xend, 
      yend = y + 0.01
    ),
    colour = "grey43",
    show.legend = FALSE
  ) +
  geom_label(
    data = stats,
    aes(
      x = phase, 
      y = y + 0.02, 
      label = pFormat
    ),
    label.size = 0,
    label.padding = unit(0.01, "lines"),
    show.legend = FALSE,
    fill = "white",
    size = 3,
    family = "Arial Unicode MS"
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Cell cycle phase",
    y = "",
    title = "Cell cycle"
  ) +
  guides(
    fill = guide_legend(title = "Condition", override.aes = list(size = 6))
  )

plotRmarkdown(p)

```

<div class="caption">
Cell cycle analysis in miR34a asRNA over-expressing PC3, Skov3, and Saos2 stable cell lines compared to their respective mock control. The points represent the values obtained from each independant experiment (PC3 n = 5; Skov3 n = 3) whereas, the gray shadow represents the distribution of those points. 95% confidence interval (vertical black lines), mean (horizontal black lines), and p-values are shown with the horizontal line under the p-value indicating the comparison that was tested.
</div>

<br></br>

####_Conclusion_

An increase in the percentage of cells in G1-phase can be seen in cell lines overexpressing miR34a asRNA. In addition, a concominant decrease of cells in G2-phase can be observed.

<br></br><br></br>