### 2017-11-14 inst/suppFigure1a_TCGAcorrelationTable/suppFigure1a_TCGAcorrelationTable.Rmd

---
title: "Supplementary Figure 1a"
subtitle: "TCGA correlation table"
author: "Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}

packages <- c(
    "tidyverse",
    "printr",
    "ggthemes",
    "readr",
    "miR34AasRNAproject",
    "grid",
    "gtable"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst/"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/data"

```

<br></br>

```{r, out.width = "850px"}

url <- "suppFigure1a_TCGAcorrelationTable/table.png"
knitr::include_graphics(file.path(projectUrl, url))

```

<div class = "caption">
R-squared and p-values from the correlation analysis investigating the correlation between miR34a and miR34a asRNA expression in p53 wild type (wt) and mutated (mut) samples within TCGA cancer types. Bladder Urothelial Carcinoma (BLCA), Breast invasive carcinoma (BRCA), Head and Neck squamous cell carcinoma (HNSC), Lower Grade Glioma (LGG), Liver hepatocellular carcinoma (LIHC), Lung adenocarcinoma (LUAD), Lung squamous cell carcinoma (LUSC), Ovarian serous cystadenocarcinoma (OV), Prostate adenocarcinoma (PRAD), Skin Cutaneous Melanoma (SKCM), Stomach adenocarcinoma (STAD).
</div>

<br></br><br></br>