### 2017-11-14 inst/suppFigure1c_polyadenylation/suppFigure1c_polyadenylation.Rmd

---
title: "Supplementary figure 1c"
subtitle: "miR34a asRNA polyadenylation"
author: "Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}

packages <- c(
    "tidyverse",
    "printr",
    "ggthemes",
    "readr",
    "miR34AasRNAproject",
    "grid",
    "gtable"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst/"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/data"

```

####_Introduction_

We desired to access the polyadenylation status of the miR34a asRNA transcript.

<br></br>

####_Methods_

<br></br>

_Cell culture and PCR_
All cell lines were cultured at 5% CO2 and 37° C with HEK293T cells cultured in DMEM high glucose (Hyclone). All growth mediums were supplemented with 10% heat-inactivated FBS and 50 μg/ml of streptomycin and 50 μg/ml of penicillin. RNA was extracted using the RNeasy mini kit (Qiagen) and subsequently treated with DNase (Ambion Turbo DNA-free, Life Technologies). 500ng RNA was used for cDNA synthesis using MuMLV (Life Technologies) and a 1:1 mix of oligo(dT) and random nanomers. PCR was then performed and the results analyzed on a 2% agarose gel.

<br></br>

_Primers_

```{r}

primers <- data.frame(
    name=c(
        "miR34aAS_F1",
        "miR34aAS_R1",
        "miR34aHG_F",
        "miR34aHG_R",
        "B-actin_F",
        "B-actin_R",
        "U48_F",
        "U48_R"
    ),
    sequence=c(
        "AGC GGC ATC TCC TCC ACC TGA AA",
        "TTG CCT CGT GAG TCC AAG GAG AAT",
        "TCT GCT CCA GTG GCT GAT GAG AAA",
        "GTT CAC TGG CCT CAA AGT TGG CAT",
        "AGG TCA TCA CCA TTG GCA ATG AG",
        "CTT TGC GGA TGT CCA CGT CA",
        "AGT GAT GAT GAC CCC AGG TA",
        "GGT CAG AGC GCT GCG GTG AT"
    )
)
primers

```

<br></br>

####_Results_

```{r, out.width = "500px"}

url <- "suppFigure1c_polyadenylation/polyadenylation.png"
knitr::include_graphics(file.path(projectUrl, url))

```

<div class = "caption">
Polyadenylation status of spliced and unspliced miR34a asRNA in HEK293T cells. 
</div>

<br></br>

####_Conclusions_

Results indicate that miR34a asRNA is polyadenylated.

<br></br><br></br>