### 2017-11-20 inst/coding_potential_cpat/coding_potential_cpat.Rmd

---
title: "Figure 1f"
subtitle: "miR34a asRNA coding potential via CPAT"
author: "Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}

packages <- c(
    "tidyverse",
    "printr",
    "ggthemes",
    "readr",
    "miR34AasRNAproject",
    "grid",
    "gtable"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst/"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/data"

```

####_Introduction_

We utilized a bioinformatic approach to evaluate the coding potential of the miR34a asRNA transcript. The Coding-potential assessment tool uses a linear regression model to evaluate coding-potential by examining ORF length, ORF coverage, Fickett score and hexamer score. We further confirmed these results using the Coding-potential Calculator which utilizes a support based machine-based classifier and accesses an alternate set of discriminatory features.

<br></br>

####_Methods_

Protein-coding capacity was evaluated using the [Coding-potential Assessment Tool](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3616698/) and [Coding-potential Calculator](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1933232/) with default settings. Transcript sequences for use with Coding-potential Assessment Tool were downloaded from the UCSC genome browser using the following IDs: HOTAIR (ENST00000455246.1), XIST (ENST00000429829.1), β-actin (ENST00000331789.5), Tubulin (ENST00000427480.1), and MYC (ENST00000377970). Transcript sequences for use with Coding-potential Calculator were downloaded from the UCSC genome browser using the following IDs: HOTAIR (uc031qho.1), β-actin (uc003soq.4).

<br></br>

####_Results_

_Coding Potential Assesment Tool_

```{r}

url <- "figure1f_codingPotentialCPAT/results.txt"
cpat <-read.table(url(file.path(projectUrl, url)), header = TRUE, sep = "\t")
cpat

```

<div class = "caption">
Coding potential analysis results from the Coding-potential Assessment Tool including miR34a asRNA and two characterized non-coding transcripts (HOTAIR and XIST) and 3 known coding transcripts (β-actin, tubulin, and MYC). 
</div>

*Results for Coding-potential Calculator are located in Supplementary figure 1g.

<br></br>

####_Conclusions_

Results indicated that miR34a asRNA has a similar lack of coding capacity to the known non-coding transcripts HOTAIR and XIST and differs greatly when examining these parameters to the known coding transcripts beta-actin, tubulin, and MYC. However, to fully evaluate coding potential methods such as mass spectrometry or ribosome profiling must be used.

<br></br><br></br>