### 2017-11-20 inst/isoforms/isoforms.Rmd

---
title: "Supplementary figure 1d"
subtitle: "miR34a asRNA isoform cloning"
author: "Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}

packages <- c(
    "tidyverse",
    "printr",
    "ggthemes",
    "readr",
    "miR34AasRNAproject",
    "grid",
    "gtable"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst/"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/data"

```

####_Introduction_

We investigated the propensity of miR34a asRNA to be alternatively spliced, using PCR cloning and sequencing.

<br></br>

####_Methods_

<br></br>

_Cell culture and cloning_

All cell lines were cultured at 5% CO2 and 37° C with U2OS cells cultured in McCoy’s 5a (Life Technologies), . All growth mediums were supplemented with 10% heat-inactivated FBS and 50 μg/ml of streptomycin and 50 μg/ml of penicillin. RNA was extracted using the RNeasy mini kit (Qiagen) and subsequently treated with DNase (Ambion Turbo DNA-free, Life Technologies). 500ng RNA was used for cDNA synthesis using MuMLV (Life Technologies) and a 1:1 mix of oligo(dT) and random nanomers. PCR was performed with the specified primer sets with annealing temperature 60˚C unless otherwise indicated. PCR was run for 35 cycles for those primer sets with annealing temperature 60˚C and 40 cycles in other cases. Resulting PCR products were run on a agarose gel, gel extracted with the QIAquick gel extraction kit (Qiagen), cloned using the Strata Cloning kit (Agilent), and sequenced. The sequencing results were visualized in the UCSC genome browser.

<br></br>

_Primers_

```{r}

primers <- data.frame(
    name=c(
        "F12",
        "R1",
        "R2",
        "R3",
        "Ex3R1"
    ),
    sequence=c(
        "AAACACAAGCGTTTACCTGGGTGC",
        "TTGCCTCGTGAGTCCAAGGAGAAT",
        "ATAGGTTCATTTGCCCGATGTGCC",
        "CCACAGCTGTTGCTTCTGAATGCT",
        "TCAGATAGGTACCAAAAATGATGGCCGCAACTAATGACGGAT"
    )
)
primers

```

<br></br>

####_Results_

```{r, out.width = "850px"}

url <- "suppFigure1d_isoforms/isoforms.png"
knitr::include_graphics(file.path(projectUrl, url))

```

<div class = "caption">
Sequencing results from the analysis of miR34a asRNA isoforms in U2OS cells. miR34a AS ref. refers to the full length transcript as defined by the 3’-RACE and primer walk assay.
</div>

<br></br>

####_Conclusions_

Results indicate that the miR34a asRNA transcript is post-transcriptionally spliced to form several different isoforms.

<br></br><br></br>