### 2017-11-20 inst/mir34a_locus/mir34a_locus.Rmd
more of the previous commit 4

---
title: "Figure 1a"
subtitle: "miR34a Locus"
author: "Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}

packages <- c(
    "tidyverse",
    "printr",
    "ggthemes",
    "readr",
    "miR34AasRNAproject",
    "grid",
    "gtable"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/"

```

####_Introduction_

An antisense transcript to the miR34a HG was detected using the UCSC genome browser.

<br></br>

####_Methods_

The USCS genome browser was utilized for the bioinformatic evaluation of antisense transcription at the miR34a locus.

<br></br>

####_Results_

```{r, out.width = "850px"}

url <- fileMap(type = "png")["figure1a"][[1]]
knitr::include_graphics(file.path(projectUrl, url))

```

<div class = "caption">
A schematic picture of the miR34a locus from the UCSC genome browser (hg38) including miR34a HG and mature miR34a, and LINC01759. H3K4me3 ChIP-seq and conservation over the locus is also shown.
</div>

<br></br>

####_Conclusion_

GENCODE annotation supports the presence of a transcript orininating in the antisense orientation from the miR34a gene.

<br></br><br></br>