### 2017-11-20 inst/coding_potential_cpc/coding_potential_cpc.Rmd
more of the previous commit 5

---
title: "Supplementary Figure 1g"
subtitle: "miR34a asRNA coding potential via CPC"
author: "Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}

packages <- c(
    "tidyverse",
    "printr",
    "ggthemes",
    "readr",
    "miR34AasRNAproject",
    "grid",
    "gtable"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/"

```

```{r}

url <- fileMap(type = "rds")["suppFigure1g"][[1]]
cpat <- read_rds(url(file.path(dataUrl, url)))
cpat

```

<div class = "caption">
Analysis of coding potential of the miR34a asRNA transcript using the Coding-potential Calculator including miR34a asRNA, the characterized non-coding transcript HOTAIR, and the known coding transcript β-actin. 
</div>

<br></br>