### 2017-11-20 inst/primer_walk_schematic/primer_walk_schematic.Rmd

---
title: "Supplementary Figure 1b"
subtitle: "Primer walk schematic"
author: "Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}

packages <- c(
    "tidyverse",
    "printr",
    "ggthemes",
    "readr",
    "miR34AasRNAproject",
    "grid",
    "gtable"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/"

```

<br></br>

```{r, out.width = "850px"}

#url <- "suppFigure1b_primerWalkSchematic/schematic_hg38.png"
#knitr::include_graphics(file.path(projectUrl, url))

url <- fileMap(type = "png")["suppFigure1b"][[1]]
knitr::include_graphics(file.path(projectUrl, url))

```

<div class = "caption">
A schematic picture of the miR34a asRNA locus including the location of the primers used in the primer walk assay.
</div>

<br></br>