### 2017-11-27 inst/cellular_localization_encode/cellular_localization_encode.Rmd
adding cellular_localization_encode

---
title: "miR34a asRNA cellular localization"
author: "Jason T. Serviss"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}

packages <- c(
    "tidyverse",
    "grid",
    "gtable",
    "miR34AasRNAproject"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/"

```

#### _Introduction_

<br></br>

#### _Methods_

<br></br>

#### _Results_

```{r}

#SK-N-SH has a larger proportion of GAPDH in the nucleus than cytoplasm
#The variation in miR34a asRNA expression is way too large for SK-MEL-5
#K562, HT1080, SK-N-DZ, and NCI-H460 have no or low miR34a asRNA expression
#use GAPDH instead of ACTB for cytoplasmic fraction, seems to be a better representation.

fdata <- data %>% 
  filter(!`Biosample term name` %in% c("K562", "HT1080", "SK-N-DZ", "NCI-H460", "SK-MEL-5", "SK-N-SH")) %>% 
  filter(gene_name != "ACTB") %>%
  select(`Biosample term name`, gene_id, gene_name, fraction, `Library made from`, `Biological replicate(s)`, TPM) %>%
  group_by(`Biosample term name`, gene_id, gene_name, fraction) %>%
  mutate(tmp = n() == 2) %>%
  filter(tmp | `Library made from` == "polyadenylated mRNA") %>%
  ungroup() %>%
  select(-tmp) %>%
  group_by(`Biosample term name`, gene_id, gene_name, `Biological replicate(s)`) %>%
  mutate(frac = TPM / sum(TPM)) %>%
  ungroup()

stats <- fdata %>%
  group_by(`Biosample term name`, gene_id, gene_name, fraction) %>%
  filter(n() == 2) %>%
  summarize(
    n = n(),
    mean = mean(frac),
    sd_min = mean(frac) + sd(frac),
    sd_max = mean(frac) - sd(frac)
  )  %>%
  ungroup()
```

```{r, fig.align='center', fig.height=8, fig.width=12, warning = FALSE}
p <- stats %>%
  ggplot(aes(gene_name, mean, color = fraction)) +
  geom_point(position = position_dodge(width = 1)) +
  facet_wrap(~`Biosample term name`, ncol = 1) +
  geom_errorbar(aes(ymax = sd_max, ymin = sd_min), width = 0, position = position_dodge(width = 1)) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  labs(
    y = "Mean %",
    x = "Gene"
  ) +
  guides(colour = guide_legend(title = "Fraction"))

plotRmarkdown(p) +
  theme(legend.position = "top")

figure <- plotPDF(p)

# path <- file.path("~/GitHub/miR34a_asRNA_project/inst", fileMap(type = "pdf")["figure2a"][[1]])
# ggsave(
#  plot = figure,
#  filename = path,
#  device = cairo_pdf,
#  height = 60,
#  width = 83,
#  units = "mm"
# )
```

<div class="caption">
Write figure legend.
</div>

<br></br>

#### _Conclusions_

<br></br><br></br>