### 2017-12-04 inst/tcga_correlation_table/tcga_correlation_table.Rmd

---
title: "Supplementary Figure 1a"
subtitle: "TCGA correlation table"
author: "Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}

packages <- c(
    "tidyverse",
    "printr",
    "ggthemes",
    "readr",
    "miR34AasRNAproject",
    "grid",
    "gtable"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/"

```

<br></br>

Note: in cases where less than 3 samples are in any group, r-estimates and p-values are not reported.

```{r, warning = FALSE}
url <- fileMap(type = "rds")["figure1c"][[1]]
data <- read_rds(gzcon(url(file.path(dataUrl, url))))

#normalize expression values
data <- mutate(
    data,
    RP3 = log2(RP3 / max(RP3)),
    miR34a = log2(miR34a / max(miR34a))  
  )

#add BRCA subtypes
data <- data %>%
  mutate(PAM50 = if_else(is.na(PAM50), "", PAM50)) %>%
  mutate(cancer_PAM50 = gsub("BRCA", "BRCA ", paste(cancer, PAM50, sep = "")))

#sort 
data <- arrange(data, cancer_PAM50, TP53, RP3)

#remove infinite data
data <- filter(data, !is.infinite(RP3) & !is.infinite(miR34a))

#diploid only
data <- filter(data, abs(RP3_cna) < 0.1)

#rename TP53
data <- mutate(data, TP53 = if_else(TP53 == 1, "TP53mut", "TP53wt"))

#diploid only correlation analysis
#calculate correlation for all samples per cancer type
all <- data %>% 
  filter(RP3_cna < 0.1) %>%
  group_by(cancer_PAM50) %>%
  summarize(
    n = n(),
    `all pvalue` = cor.test(RP3, miR34a, method = "spearman")$p.value,
    `all rho` = cor.test(RP3, miR34a, method = "spearman")$estimate
  ) %>%
  mutate(
    `all pvalue` = format(`all pvalue`, scientific = TRUE, digits = 3),
    `all rho` = format(`all rho`, scientific = TRUE, digits = 3)
  )

#calculate correlation for p53 wt and mut samples per cancer type
p53 <- data %>%
  group_by(cancer_PAM50, TP53) %>%
  filter(n() > 3) %>%
  summarize(
    n = n(),
    pvalue = cor.test(RP3, miR34a, method = "spearman")$p.value,
    rho = cor.test(RP3, miR34a, method = "spearman")$estimate
  ) %>%
  ungroup() %>%
  mutate(
    pvalue = format(pvalue, scientific = TRUE, digits = 3),
    rho = format(rho, scientific = TRUE, digits = 3)
  ) %>%
  gather(metric, value, -(cancer_PAM50:TP53)) %>%
  unite(temp, TP53, metric, sep = " ") %>%
  spread(temp, value)

table <- full_join(all, p53, by = "cancer_PAM50") %>%
  rename(cancer = cancer_PAM50, `all n` = n) %>%
  select(
    cancer, `all n`, `all rho`, `all pvalue`, `TP53wt n`, 
    `TP53wt rho`, `TP53wt pvalue`, `TP53mut n`, `TP53mut rho`, `TP53mut pvalue`
  )

table

path <- file.path("~/GitHub/miR34a_asRNA_project/inst", fileMap(type = "pdf")["suppFigure1a"][[1]])
pdf(file = path, width = 14)
gridExtra::grid.table(table, rows = NULL)
invisible(dev.off())
```

<div class = "caption">
R-squared and p-values from the correlation analysis investigating the correlation between miR34a and miR34a asRNA expression in p53 wild type (wt) and mutated (mut) samples within TCGA cancer types. Bladder Urothelial Carcinoma (BLCA), Breast invasive carcinoma (BRCA), Head and Neck squamous cell carcinoma (HNSC), Lower Grade Glioma (LGG), Liver hepatocellular carcinoma (LIHC), Lung adenocarcinoma (LUAD), Lung squamous cell carcinoma (LUSC), Ovarian serous cystadenocarcinoma (OV), Prostate adenocarcinoma (PRAD), Skin Cutaneous Melanoma (SKCM), Stomach adenocarcinoma (STAD).
</div>

<br></br><br></br>

```{r}
sessionInfo()
```