### 2017-12-10 inst/isoforms/isoforms.Rmd
added primer combos and annealing info to isoforms

---
title: "Figure 1-Supplement 2c"
subtitle: "miR34a asRNA isoform cloning"
author: "Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}

packages <- c(
    "tidyverse",
    "printr",
    "ggthemes",
    "readr",
    "miR34AasRNAproject",
    "grid",
    "gtable"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/"

```

#### _Introduction_

We investigated the propensity of miR34a asRNA to be alternatively spliced, using PCR cloning and sequencing.

<br></br>

#### _Methods_

<br></br>

_Cell culture and cloning_

All cell lines were cultured at 5% CO2 and 37° C with U2OS cells cultured in McCoy’s 5a (Life Technologies), . All growth mediums were supplemented with 10% heat-inactivated FBS and 50 μg/ml of streptomycin and 50 μg/ml of penicillin. RNA was extracted using the RNeasy mini kit (Qiagen) and subsequently treated with DNase (Ambion Turbo DNA-free, Life Technologies). 500ng RNA was used for cDNA synthesis using MuMLV (Life Technologies) and a 1:1 mix of oligo(dT) and random nanomers. PCR was performed with the specified primer sets with annealing temperature 60˚C unless otherwise indicated (see below). PCR was run for 35 cycles for those primer sets with annealing temperature 60˚C and 40 cycles in other cases. Resulting PCR products were run on a agarose gel, gel extracted (8 in total) with the QIAquick gel extraction kit (Qiagen), cloned using the Strata Cloning kit (Agilent), and sequenced. The sequencing results were visualized in the UCSC genome browser. Results aligning to the opposite strand and spurious results (several small 24bp alignments) were removed from the final figure.

<br></br>

_Primers_

```{r}

primers <- data.frame(
    name=c(
        "miR34a AS F12",
        "miR34a AS R1",
        "miR34a AS R2",
        "miR34a AS R3",
        "miR34a AS Ex3 R1"
    ),
    sequence=c(
        "AAACACAAGCGTTTACCTGGGTGC",
        "TTGCCTCGTGAGTCCAAGGAGAAT",
        "ATAGGTTCATTTGCCCGATGTGCC",
        "CCACAGCTGTTGCTTCTGAATGCT",
        "TCAGATAGGTACCAAAAATGATGGCCGCAACTAATGACGGAT"
    )
)
primers

```

Primer combinations:
miR34a AS F12 + miR34a AS R1
miR34a AS F12 + miR34a AS R2
miR34a AS F12 + miR34a AS R3
miR34a AS F12 + miR34a AS Ex3 R1 (56 degrees annealing)
miR34a AS F12 + miR34a AS Ex3 R1 (58 degrees annealing)

<br></br>

#### _Results_

```{r, out.width = "850px"}

url <- fileMap(type = "png")["Figure 1-Supplement 2c"][[1]]
knitr::include_graphics(file.path(projectUrl, url))

```

<div class = "caption">
Sequencing results from the analysis of miR34a asRNA isoforms in U2OS cells. miR34a AS ref. refers to the full length transcript as defined by the 3’-RACE and primer walk assay.
</div>

<br></br>

#### _Conclusions_

Results indicate that the miR34a asRNA transcript is post-transcriptionally spliced to form several different isoforms.

<br></br><br></br>

```{r}
sessionInfo()
```