### 2018-02-27 inst/tcga_survival/fiveYR.Rmd
adding further survival functions and survivalKM analysis

---
title: "Figure 4b"
subtitle: "5 year survival analysis"
author: "Jimmy Van den Eynden and Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}
packages <- c(
  "tidyverse",
  "printr",
  "ggthemes",
  "readr",
  "miR34AasRNAproject",
  "grid",
  "gtable",
  "survival",
  "KMsurv"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/"

```

#### _Introduction_

#### _Methods_

<br></br>

#### _Results_
```{r}

url <- fileMap(type = "rds")["Figure 4b"][[1]]
data <- read_rds(gzcon(url(file.path(dataUrl, url))))

```

### 5 year survival analysis

```{r}
correlation_matrix <- filter(data, !is.na(TP53) & !is.na(lncTAM34a))
cancers <- sort(unique(correlation_matrix$Cancer))

#TP53 status independant
#Set up all combinations of variables to calcualte survival probability for
independant <- expand.grid(
  gene = c("lncTAM34a", "miR34a", "TP53"),
  cancer = cancers,
  geneLevel = c("low", "others")
) %>% 
  as_tibble() %>%
  arrange(cancer, gene, geneLevel) %>%
  mutate_if(is.factor, as.character) %>%
  group_by(gene, cancer, geneLevel) %>%
  #add data of the corresponding cancer
  mutate(data = map(cancer, function(c, cm = correlation_matrix) {
    filter(cm, Cancer == c)
  })) %>%
  #select the gene to be examined
  mutate(data = pmap(list(gene, data), function(g, d) {
      select(d, g, vitalStatus, FU)
  })) %>%
  #calculate the percentile for the gene's expression for each sample
  mutate(data = map2(gene, data, function(g, d) {
    if(any(grepl("TP53", colnames(d)))) {
      mutate(d, percentile = NA)
    } else {
      mutate(d, percentile = ntile(x = d[[1]], n = 10))
    }
  })) %>%
  #classify 10th percentile and under as "low" and the rest as "others"
  #classify TP53 mutated as low and WT as others
  mutate(data = map(data, function(d) {
    if(any(grepl("TP53", colnames(d)))) {
      mutate(d, class = if_else(TP53, "low", "others"))
    } else {
     mutate(d, class = if_else(percentile == 1, "low", "others")) 
    }
  })) %>%
  #filter data class to correspond to expected geneLevel
  mutate(data = map2(geneLevel, data, function(gl, d) {
    filter(d, class == gl)
  })) %>%
  #calculate survival probability
  mutate(surv.obj = map(data, ~get_survival(.x))) %>%
  mutate(surv.fit = map(surv.obj, ~survival_fit(.x))) %>%
  mutate(surv.prob = map_dbl(surv.fit, ~calcSurvProb(.x))) %>%
  ungroup() %>%
  select(-data, -surv.fit) %>%
  mutate(type = "independant")

# TP53 status dependent
#Set up all combinations of variables to calcualte survival probability for
dependant <- expand.grid(
  gene = c("lncTAM34a", "miR34a"),
  mutation = c(TRUE, FALSE),
  cancer = cancers,
  geneLevel = c("low", "others")
) %>% 
  as_tibble() %>%
  arrange(cancer, mutation, gene, geneLevel) %>%
  mutate_if(is.factor, as.character) %>%
  group_by(gene, mutation, cancer, geneLevel) %>%
  #add data of the corresponding cancer and TP53 status
  mutate(data = pmap(list(cancer, mutation), function(c, m, cm = correlation_matrix) {
    filter(cm, Cancer == c & TP53 == m)
  })) %>%
  #select the gene to be examined
  mutate(data = pmap(list(gene, data), function(g, d) {
      select(d, g, vitalStatus, FU)
  })) %>%
  #calculate the percentile for the gene's expression for each sample
  mutate(data = map2(gene, data, function(g, d) {
    mutate(d, percentile = ntile(x = d[[1]], n = 10))
  })) %>%
  #classify 10th percentile and under as "low" and the rest as "others"
  mutate(data = map(data, function(d) {
    mutate(d, class = if_else(percentile == 1, "low", "others"))
  })) %>%
  #filter data class to correspond to expected geneLevel
  mutate(data = map2(geneLevel, data, function(gl, d) {
    filter(d, class == gl)
  })) %>%
  #remove comparisons where there is not enough data
  mutate(bool = map_lgl(data, function(d) {
    !nrow(d) <= 1
  })) %>%
  filter(bool) %>%
  #calculate survival probability
  mutate(surv.obj = map(data, ~get_survival(.x))) %>%
  mutate(surv.fit = map(surv.obj, ~survival_fit(.x))) %>%
  mutate(surv.prob = map_dbl(surv.fit, ~calcSurvProb(.x))) %>%
  ungroup() %>%
  select(-(data:surv.fit)) %>%
  mutate(type = "dependent")

#combine the analyses and filter those to be included in the figure
survData <- bind_rows(dependant, independant) %>%
  filter(!is.na(surv.prob)) %>%
  filter((type == "dependent") | (type == "independant" & gene == "TP53")) %>%
  filter(is.na(mutation) | mutation == FALSE) %>%
  group_by(gene, mutation, cancer, type) %>%
  mutate(n = n()) %>%
  filter(n == 2) %>%
  summarize(relSurv = surv.prob[geneLevel == "low"] - surv.prob[geneLevel == "others"]) %>%
  ungroup() %>%
  select(-mutation, -type) %>%
  spread(gene, relSurv)
```

```{r, warning = FALSE}
#Add the colors for the plot to the data
colors <- getCancerColors()
survData <- full_join(survData, colors, by = "cancer")

#add ggplot theme
theme_remove_all <- theme_remove_all()

#plot 1
p1 <- mainPlot(survData, "TP53", "lncTAM34a", tRA = theme_remove_all)
p2 <- hPlot(survData, "TP53", tRA = theme_remove_all)
renderPlots(p1, p2)

#plot 2
p1 <- mainPlot(survData, "lncTAM34a", "miR34a", tRA = theme_remove_all)
p2 <- hPlot(survData, "lncTAM34a", tRA = theme_remove_all)
renderPlots(p1, p2)

#plot 3
p1 <- mainPlot(survData, "miR34a", "TP53", tRA = theme_remove_all)
p2 <- hPlot(survData, "miR34a", tRA = theme_remove_all)
renderPlots(p1, p2)
```