### 2018-02-28 inst/tcga_survival_total/tcga_survival_total.Rmd
last updates before CDD submission

---
title: "Figure 4a and Supplementary Figure 5"
subtitle: "Survival curves"
author: "Jimmy Van den Eynden and Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

<style>
.col2 { 
    columns: 2 200px; /* number of columns and width in pixels*/ 
    -webkit-columns: 2 200px; /* chrome, safari */ 
    -moz-columns: 2 200px; /* firefox */ 
}
</style>

```{r, echo = FALSE, message = FALSE}
packages <- c(
  "tidyverse",
  "printr",
  "ggthemes",
  "readr",
  "miR34AasRNAproject"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/"

```

#### _Introduction_
Kaplan-Meier survival curves comparing the effects of TP53-mutated samples,
low lncTAM34a expression and low miR34a expression to control samples in
17 cancers from TCGA. Panels showing survival curves based on gene expression
include only TP53 wild type patients where RNAseq data exists.

#### _Methods_
RNAseq data was downloaded from TCGA and processed as described previously. 
Briefly, RNAseq data were aligned to the human hg19 assembly and quantified 
using GENCODE (v19) annotated HTSeq-counts and FPKM normalizations. Expression 
data from miR34a and lncTAM34a (identified as RP3-510D11.2) were used for 
further analysis.

Survival analysis was performed on TCGA vital state and follow-up data,
downloaded from GDC on 27/10/2017 using the R survival package.

<br></br>

#### _Results_
```{r}

url <- fileMap(type = "rds")["Figure 4b"][[1]]
data <- read_rds(gzcon(url(file.path(dataUrl, url))))

```

```{r}
#TP53 status independant
#Set up all combinations of variables to calcualte survival probability for
independant <- expand.grid(
  gene = c("lncTAM34a", "miR34a", "TP53"),
  cancer = sort(unique(data$Cancer))
) %>% 
  as_tibble() %>%
  mutate_if(is.factor, as.character) %>%
  group_by(gene, cancer) %>%
  #add data of the corresponding cancer and, for gene expression, filter WT TP53
  mutate(data = map2(cancer, gene, function(c, g, cm = data) {
    if(g == "TP53") {
      filter(cm, Cancer == c)
    } else {
      filter(cm, Cancer == c & TP53 == FALSE)
    }
  })) %>%
  #select the gene to be examined
  mutate(data = pmap(list(gene, data), function(g, d) {
      select(d, g, vitalStatus, FU)
  })) %>%
  #remove samples where gene expression/mutation status is na
  mutate(data = map(data, function(x) filter(x, is.na(x[[1]]) == FALSE))) %>%
  #remove cancers where comparisons without data exist
  mutate(bool = map_lgl(data, function(x) nrow(x) == 0)) %>%
  group_by(cancer) %>%
  filter(!any(bool)) %>%
  #calculate the percentile for the gene's expression for each sample
  mutate(data = map2(gene, data, function(g, d) {
    if(any(grepl("TP53", colnames(d)))) {
      mutate(d, percentile = NA)
    } else {
      mutate(d, percentile = ntile(x = d[[1]], n = 10))
    }
  })) %>%
  #classify 10th percentile and under as "low" and the rest as "others"
  #classify TP53 mutated as low and WT as others
  mutate(data = map(data, function(d) {
    if(any(grepl("TP53", colnames(d)))) {
      mutate(d, class = if_else(TP53, "low", "others"))
    } else {
     mutate(d, class = if_else(percentile == 1, "low", "others")) 
    }
  })) %>%
  #calculate survival probability
  mutate(surv.obj = map(data, ~get_survival(.x))) %>%
  mutate(surv.fit = map2(data, surv.obj, function(x, y) survival_fit(y, x$class))) %>%
  mutate(p.value = map2_dbl(data, surv.obj, function(x, y) survival_p(y, x$class))) %>%
  ungroup()
```

<div class="col2">
```{r}
trash <- pmap(
  list(independant$data, independant$surv.fit, independant$gene, independant$cancer, independant$p.value), 
  function(v, w, x, y, z) {
    plotKM(v, w, x, y, z)
  }
)
```
</div>