### 2018-03-05 inst/tcga_survival/tcga_survival.Rmd

---
title: "Figure 4b"
subtitle: "5 year survival analysis"
author: "Jimmy Van den Eynden and Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}
packages <- c(
  "tidyverse",
  "printr",
  "ggthemes",
  "readr",
  "miR34AasRNAproject",
  "grid",
  "gtable",
  "survival",
  "KMsurv"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

<br></br>

#### _Introduction_
Correlation analysis between the effects on the 5-year survival
probability of TP53-mutated samples, low lncTAM34a expression, and low
miR34a expression as indicated. For each variable the 5-year survival
probability was compared to the control group (negative values indicate lower
survival, positive values indicate higher survival). Spearman correlation
coefficients are given on top left of each plot. Each dot indicates one cancer
type. Boxplots on the bottom summarize the effects for
the parameter on the x-axis, with indication of P values, as calculated using
paired Wilcoxon signed rank test. Low expression was defined as TP53 non-
mutated samples having expression values in the bottom 10th percentile.

<br></br>

#### _Methods_
RNAseq data was downloaded from TCGA and processed as described previously. 
Briefly, RNAseq data were aligned to the human hg19 assembly and quantified 
using GENCODE (v19) annotated HTSeq-counts and FPKM normalizations. Expression 
data from miR34a and lncTAM34a (identified as RP3-510D11.2) were used for 
further analysis.

Survival analysis was performed on TCGA vital state and follow-up data,
downloaded from GDC on 27/10/2017 using the R survival package.

<br></br>

#### _Results_

```{r}

data <- getData("Figure 4b")
correlation_matrix <- filter(data, !is.na(TP53) & !is.na(lncTAM34a))
cancers <- sort(unique(correlation_matrix$Cancer))

#TP53 status independant
#Set up all combinations of variables to calcualte survival probability for
independant <- expand.grid(
  gene = c("lncTAM34a", "miR34a", "TP53"),
  cancer = cancers,
  geneLevel = c("low", "others")
) %>% 
  as_tibble() %>%
  arrange(cancer, gene, geneLevel) %>%
  mutate_if(is.factor, as.character) %>%
  group_by(gene, cancer, geneLevel) %>%
  #add data of the corresponding cancer
  mutate(data = map(cancer, function(c, cm = correlation_matrix) {
    filter(cm, Cancer == c)
  })) %>%
  #select the gene to be examined
  mutate(data = pmap(list(gene, data), function(g, d) {
      select(d, g, vitalStatus, FU)
  })) %>%
  #calculate the percentile for the gene's expression for each sample
  mutate(data = map2(gene, data, function(g, d) {
    if(any(grepl("TP53", colnames(d)))) {
      mutate(d, percentile = NA)
    } else {
      mutate(d, percentile = ntile(x = d[[1]], n = 10))
    }
  })) %>%
  #classify 10th percentile and under as "low" and the rest as "others"
  #classify TP53 mutated as low and WT as others
  mutate(data = map(data, function(d) {
    if(any(grepl("TP53", colnames(d)))) {
      mutate(d, class = if_else(TP53, "low", "others"))
    } else {
     mutate(d, class = if_else(percentile == 1, "low", "others")) 
    }
  })) %>%
  #filter data class to correspond to expected geneLevel
  mutate(data = map2(geneLevel, data, function(gl, d) {
    filter(d, class == gl)
  })) %>%
  #calculate survival probability
  mutate(surv.obj = map(data, ~get_survival(.x))) %>%
  mutate(surv.fit = map(surv.obj, ~survival_fit(.x))) %>%
  mutate(surv.prob = map_dbl(surv.fit, ~calcSurvProb(.x))) %>%
  ungroup() %>%
  select(-data, -surv.fit) %>%
  mutate(type = "independant")

# TP53 status dependent
#Set up all combinations of variables to calcualte survival probability for
dependant <- expand.grid(
  gene = c("lncTAM34a", "miR34a"),
  mutation = c(TRUE, FALSE),
  cancer = cancers,
  geneLevel = c("low", "others")
) %>% 
  as_tibble() %>%
  arrange(cancer, mutation, gene, geneLevel) %>%
  mutate_if(is.factor, as.character) %>%
  group_by(gene, mutation, cancer, geneLevel) %>%
  #add data of the corresponding cancer and TP53 status
  mutate(data = pmap(list(cancer, mutation), function(c, m, cm = correlation_matrix) {
    filter(cm, Cancer == c & TP53 == m)
  })) %>%
  #select the gene to be examined
  mutate(data = pmap(list(gene, data), function(g, d) {
      select(d, g, vitalStatus, FU)
  })) %>%
  #calculate the percentile for the gene's expression for each sample
  mutate(data = map2(gene, data, function(g, d) {
    mutate(d, percentile = ntile(x = d[[1]], n = 10))
  })) %>%
  #classify 10th percentile and under as "low" and the rest as "others"
  mutate(data = map(data, function(d) {
    mutate(d, class = if_else(percentile == 1, "low", "others"))
  })) %>%
  #filter data class to correspond to expected geneLevel
  mutate(data = map2(geneLevel, data, function(gl, d) {
    filter(d, class == gl)
  })) %>%
  #remove comparisons where there is not enough data
  mutate(bool = map_lgl(data, function(d) {
    !nrow(d) <= 1
  })) %>%
  filter(bool) %>%
  #calculate survival probability
  mutate(surv.obj = map(data, ~get_survival(.x))) %>%
  mutate(surv.fit = map(surv.obj, ~survival_fit(.x))) %>%
  mutate(surv.prob = map_dbl(surv.fit, ~calcSurvProb(.x))) %>%
  ungroup() %>%
  select(-(data:surv.fit)) %>%
  mutate(type = "dependent")

#combine the analyses and filter those to be included in the figure
survData <- bind_rows(dependant, independant) %>%
  filter(!is.na(surv.prob)) %>%
  filter((type == "dependent") | (type == "independant" & gene == "TP53")) %>%
  filter(is.na(mutation) | mutation == FALSE) %>%
  group_by(gene, mutation, cancer, type) %>%
  mutate(n = n()) %>%
  filter(n == 2) %>%
  summarize(relSurv = surv.prob[geneLevel == "low"] - surv.prob[geneLevel == "others"]) %>%
  ungroup() %>%
  select(-mutation, -type) %>%
  spread(gene, relSurv)
```

```{r, warning = FALSE, fig.align='center', fig.height=8, fig.width=12}
#Add the colors for the plot to the data
colors <- getCancerColors()
survData <- full_join(survData, colors, by = "cancer")

#add ggplot theme
theme_remove_all <- theme_remove_all()

#plot 1
p1 <- mainPlot(survData, "TP53", "lncTAM34a", tRA = theme_remove_all)
p2 <- hPlot(survData, "TP53", tRA = theme_remove_all)
renderPlots(p1, p2)

#plot 2
p1 <- mainPlot(survData, "lncTAM34a", "miR34a", tRA = theme_remove_all)
p2 <- hPlot(survData, "lncTAM34a", tRA = theme_remove_all)
renderPlots(p1, p2)

#plot 3
p1 <- mainPlot(survData, "miR34a", "TP53", tRA = theme_remove_all)
p2 <- hPlot(survData, "miR34a", tRA = theme_remove_all)
renderPlots(p1, p2)
```