### 2018-05-16 inst/supp_doc2/supp_doc2.Rmd
adding rmd for supp doc 2

---
title: "Supplementary Document 2"
output: html_document
---

## Figure 1b
###

| name             | sequence                         |
| ---------------- |:--------------------------------:|
| miR34a asRNA F1  | AGC GGC ATC TCC TCC ACC TGA AA   |
| miR34a asRNA R1  | TTG CCT CGT GAG TCC AAG GAG AAT  |
| miR34a HG F      | TCT GCT CCA GTG GCT GAT GAG AAA  |
| miR34a HG R      | GTT CAC TGG CCT CAA AGT TGG CAT  |
| ß-actin Fwd      | AGG TCA TCA CCA TTG GCA ATG AG   |
| ß-actin Rev      | CTT TGC GGA TGT CCA CGT CA       |

## Figure 1d

| name              | sequence                                                                                            |
| ----------------- |:---------------------------------------------------------------------------------------------------:|
| miR34a asRNA F10  | ACG CGT CTC TCC AGC CCG GGA T                                                                       |
| polyT T7 FAM      | CAG TGA ATT GTA ATA CGA CTC ACT ATA GGG ACA TCC GTA GCT CGT CCA GGA CCC TTT TTT TTT TTT TTT TTT VN  |
| miR34a asRNA F1   | AGC GGC ATC TCC TCC ACC TGA AA                                                                      |
| FAM primer        | CCG TAG CTC GTC CAG GAC CC                                                                          |

## Figure 2a

| name          | sequence                         |
| ------------- |:--------------------------------:|
| ß-actin Fwd   | AGG TCA TCA CCA TTG GCA ATG AG   |
| ß-actin Rev   | CTT TCG GGA TGT CCA CGT CA       |
| miR34a HG F   | TCT GCT CCA GTG GCT GAT GAG AAA  |
| miR34a HG R   | GTT CAC TGG CCT CAA AGT TGG CAT  |
| miR34a AS F1  | AGC GGC ATC TCC TCC ACC TGA AA   |
| miR34a AS R1  | TTG CCT CGT GAG TCC AAG GAG AAT  |

## Figure 2b

| name          | sequence                         |
| ------------- |:--------------------------------:|
| ß-actin Fwd   | AGG TCA TCA CCA TTG GCA ATG AG   |
| ß-actin Rev   | CTT TCG GGA TGT CCA CGT CA       |
| miR34a HG F   | TCT GCT CCA GTG GCT GAT GAG AAA  |
| miR34a HG R   | GTT CAC TGG CCT CAA AGT TGG CAT  |
| miR34a AS F1  | AGC GGC ATC TCC TCC ACC TGA AA   |
| miR34a AS R1  | TTG CCT CGT GAG TCC AAG GAG AAT  |

## Figure 2c

**QPCR primers**

| name          | sequence                         |
| ------------- |:--------------------------------:|
| ß-actin Fwd   | AGG TCA TCA CCA TTG GCA ATG AG   |
| ß-actin Rev   | CTT TCG GGA TGT CCA CGT CA       |
| miR34a HG F   | TCT GCT CCA GTG GCT GAT GAG AAA  |
| miR34a HG R   | GTT CAC TGG CCT CAA AGT TGG CAT  |
| miR34a AS F1  | AGC GGC ATC TCC TCC ACC TGA AA   |
| miR34a AS R1  | TTG CCT CGT GAG TCC AAG GAG AAT  |

## Figure 2d

| name               | sequence                        |
| ------------------ |:-------------------------------:|
| Luc setII F        | AAG ATT CAA AGT GCG CTG CTG     |
| Luc setII R        | TTG CCT GAT ACC TGG CAG ATG     |
| Renilla pBiDir F1  | TAA CGC GGC CTC TTC TTA TTT     |
| Renilla pBiDir R1  | GAT TTG CCT GAT TTG CCC ATA     |
| ß-actin Fwd        | AGG TCA TCA CCA TTG GCA ATG AG  |
| ß-actin Rev        | CTT TGC GGA TGT CCA CGT CA      |

## Figure 2e

| name               | sequence                        |
| ------------------ |:-------------------------------:|
| Luc setII F        | AAG ATT CAA AGT GCG CTG CTG     |
| Luc setII R        | TTG CCT GAT ACC TGG CAG ATG     |
| Renilla pBiDir F1  | TAA CGC GGC CTC TTC TTA TTT     |
| Renilla pBiDir R1  | GAT TTG CCT GAT TTG CCC ATA     |
| ß-actin Fwd        | AGG TCA TCA CCA TTG GCA ATG AG  |
| ß-actin Rev        | CTT TGC GGA TGT CCA CGT CA      |

## Figure 3a

**Cloning primers**

| name                    | sequence                         |
| ----------------------- |:--------------------------------:|
| miR34aAS cloning F4     | ACG CGT CTC TCC AGC CCG GGA T    |
| miR34aAS cloning Ex3_1  | AAT GAT GGC CGC AAC TAA TGA CGG  |

**QPCR primers**

| name         | sequence                         |
| ------------ |:--------------------------------:|
| ß-actin Fwd  | AGG TCA TCA CCA TTG GCA ATG AG   |
| ß-actin Rev  | CTT TCG GGA TGT CCA CGT CA       |
| miR34a AS F1 |  AGC GGC ATC TCC TCC ACC TGA AA  |
| miR34a AS R1 |  TTG CCT CGT GAG TCC AAG GAG AAT |

## Figure 3d

| name            | sequence                         |
| --------------- |:--------------------------------:|
| miR34a ChIP F1  | AAA GTT TGC AAA GAA GGA GGC GGG  |
| miR34a ChIP R1  | AGG GAA GAA AGA ACT AGC CGA GCA  |

## Supplementary Figure 2a

| name           | sequence                         |
| -------------- |:--------------------------------:|
| miR34a AS F10  | ACG CGT CTC TCC AGC CCG GGA T    |
| miR34a AS F11  | ATC TGC GTG GTC ACC GAG AAG CA   |
| miR34a AS F12  | CGC ACG GAC TGA GAA ACA CAA G    |
| miR34a AS F13  | ACG GAG GCT ACA CAA TTG AAC AGG  |
| miR34a AS F14  | AGG GAA GAA AGA ACT AGC CGA GCA  |
| miR34a AS F15  | CAT TTG CTG CAA TAT CAC CGT GGC  |
| miR34a AS R1   | TTG CCT CGT GAG TCC AAG GAG AAT  |

## Supplementary Figure 2b

| name               | sequence                         |
| ------------------ |:--------------------------------:|
| miR34a AS F1       | AGC GGC ATC TCC TCC ACC TGA AA   |
| miR34a AS R1       | TTG CCT CGT GAG TCC AAG GAG AAT  |
| miR34a AS int1 R1  | TGC GCA AAC TAC GCG CTC T        |
| miR34a HG F        | TCT GCT CCA GTG GCT GAT GAG AAA  |
| miR34a HG R        | GTT CAC TGG CCT CAA AGT TGG CAT  |
| ß-actin Fwd        | AGG TCA TCA CCA TTG GCA ATG AG   |
| ß-actin Rev        | CTT TGC GGA TGT CCA CGT CA       |
| U48 F              | AGT GAT GAT GAC CCC AGG TA       |
| U48 R              | GGT CAG AGC GCT GCG GTG AT       |

## Supplementary Figure 2c

| name              | sequence                         |
| ----------------- |:--------------------------------:|
| miR34a AS F12     | AAA CAC AAG CGT TTA CCT GGG TGC  |
| miR34a AS R1      | TTG CCT CGT GAG TCC AAG GAG AAT  |
| miR34a AS R2      | ATA GGT TCA TTT GCC CGA TGT GCC  |
| miR34a AS R3      | CCA CAG CTG TTG CTT CTG AAT GCT  |
| miR34a AS Ex3 R1  | TGA TGG CCG CAA CTA ATG ACG GAT  |

## Supplementary Figure 3c

| name               | sequence                        |
| ------------------ |:-------------------------------:|
| Luc setII F        | AAG ATT CAA AGT GCG CTG CTG     |
| Luc setII R        | TTG CCT GAT ACC TGG CAG ATG     |
| Renilla pBiDir F1  | TAA CGC GGC CTC TTC TTA TTT     |
| Renilla pBiDir R1  | GAT TTG CCT GAT TTG CCC ATA     |
| ß-actin Fwd        | AGG TCA TCA CCA TTG GCA ATG AG  |
| ß-actin Rev        | CTT TGC GGA TGT CCA CGT CA      |

## Supplementary Figure 4a

**Cloning primers**

| name                    | sequence                         |
| ----------------------- |:--------------------------------:|
| miR34aAS cloning F4     | ACG CGT CTC TCC AGC CCG GGA T    |
| miR34aAS cloning Ex3_1  | AAT GAT GGC CGC AAC TAA TGA CGG  |

**QPCR primers**

| name          | sequence                         |
| ------------- |:--------------------------------:|
| ß-actin Fwd   | AGG TCA TCA CCA TTG GCA ATG AG   |
| ß-actin Rev   | CTT TCG GGA TGT CCA CGT CA       |
| miR34a AS F1  | AGC GGC ATC TCC TCC ACC TGA AA   |
| miR34a AS R1  | TTG CCT CGT GAG TCC AAG GAG AAT  |

## Supplementary Figure 4c

| name         | sequence                        |
| ------------ |:-------------------------------:|
| CCND1 Fwd    | CGT GGC CTC TAA GAT GAA GG      |
| CCND1 Rev    | CTG GCA TTT TGG AGA GGA AG      |
| ß-actin Fwd  | AGG TCA TCA CCA TTG GCA ATG AG  |
| ß-actin Rev  | CTT TGC GGA TGT CCA CGT CA      |