### 2018-05-16 inst/figure3c/figure3c.Rmd
updating fig3c to use polynomial regression

---
title: "Figure 3c"
subtitle: "Monitoring growth of miR34a asRNA overexpressings cells under conditions of starvation"
author: "Jason T. Serviss"
date: "31/01/2017"
output:
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

<style> 
.caption { 
  color: Black; 
  font-family: "Times New Roman", Times, serif;
  font-size: 1.0em; 
}
</style>

```{r, echo = FALSE, message = FALSE}

packages <- c(
    "tidyverse",
    "printr",
    "ggthemes",
    "readr",
    "miR34AasRNAproject",
    "grid",
    "gtable",
    "broom"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)

```

```{r}

projectUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/inst"
dataUrl <- "https://github.com/GranderLab/miR34a_asRNA_project/raw/master/"

```

<br></br>

#### _Introduction_

As p53 is a well-known regulator of senescence, growth, and apoptosis and has been shown to specifically important under conditions of cellular stress, such as starvation. Therefore, we investigated whether overexpression of miR34a AS affects growth rate under normal and starvation conditions by measuring changes in confluency over time.

<br></br>

#### _Methods_

<br></br>

_Cell culture and confluency quantification_

10^4 PC3 cells, either miR34a asRNA overexpressing or WT, were seeded in 96 well plates. After attachment (3-4h) media was replaced with either RPMI (Gibco, life technology) (supplemented with 2mM L-glutamine, 50ug/ml Penicillin-Streptomycin and 10% Fetal Calf Serum) or HBSS. Cells were incubated in Spark Multimode Microplate reader for 48 hours at 37°C with 5% CO2 in a humidity chamber. Confluency was measured every hour.

<br></br>

_Analysis_

Three technical triplicates and three biological replicates were included in each experiment and these were independantly normalized for each condition to the 0 time point afterwhich the mean was calculated for the technical replicates. A polynomial regression model was then implemented separatley for each culture condition. with the mean of the technical replicates as the dependant variable and time and cell line as the independant variables.

<br></br>

#### _Results_

_Example confluence measurment_

```{r, out.width = "450px"}

url <- fileMap(type = "png")["Supplementary Figure 5a"][[1]]
knitr::include_graphics(file.path(projectUrl, url))

```

<div class = "caption">
Representative pictures from the growth analysis shown in Figure 3C. The green color indicates the surface covered by cells as detected by the analysis software.
</div>

<br></br>

_Example phase contrast_

```{r, out.width = "450px"}

url <- fileMap(type = "png")["Supplementary Figure 5b"][[1]]
knitr::include_graphics(file.path(projectUrl, url))

```

<div class = "caption">
A representative phase contrast picture from the 12 hour time point of the growth analysis in Figure 3C.
</div>

<br></br>

```{r, message = FALSE}

data <- getData("Figure 3c")

#normalize to first time point for control (Mock)
.normFun <- function(
  data,
  Confluency,
  Treatment,
  Biological.Replicate,
  Technical.Replicate,
  Cell.line
){
  bool1 <- data$Treatment == Treatment
  bool2 <- data$`Biological Replicate` == Biological.Replicate
  bool3 <- data$`Technical Replicate` == Technical.Replicate
  bool4 <- data$Time == 0
  bool5 <- data$`Cell line` == Cell.line
  bool <- bool1 & bool2 & bool3 & bool4 & bool5
  Confluency / pull(data, Confluency)[bool]
}

data <- data %>%
  group_by(Treatment, `Biological Replicate`, `Technical Replicate`, Time, `Cell line`) %>%
  mutate(
    normConfluency = .normFun(
      data,
      Confluency,
      Treatment,
      `Biological Replicate`,
      `Technical Replicate`,
      `Cell line`
    )
  ) %>%
  ungroup()

#calculate mean for each technical replicate
data <- data %>%
  group_by(Treatment, `Biological Replicate`, Time, `Cell line`) %>%
  summarize(
    techMean = mean(normConfluency)
  ) %>%
  ungroup()

#linear model
model.data <-  data %>%
  mutate(Time = as.numeric(Time)) %>%
  group_by(Treatment) %>%
  do(glance(lm(techMean ~ poly(Time, 24) +`Cell line`, data = .))) %>%
  ungroup() %>%
  rename(pValue = p.value) %>%
  pFormat(.) %>%
  select(-pValue) %>%
  rename(p.value = pFormat)

as.data.frame(model.data)

growth.lm <- data %>%
  mutate(Time = as.numeric(Time)) %>%
  group_by(Treatment) %>%
  do(tidy(lm(techMean ~ poly(Time, 24) +`Cell line`, data = .))) %>%
  ungroup() %>%
  rename(pValue = p.value) %>%
  pFormat(.) %>%
  select(-pValue) %>%
  rename(p.value = pFormat)

as.data.frame(growth.lm)

```

```{r, fig.align='center', fig.height=8, fig.width=12}

p <- ggplot(data = NULL) +
  geom_smooth(
    data = data,
    aes(
      x = Time,
      y = techMean,
      group = `Cell line`,
      colour = `Cell line`
    ),
    method = "lm", 
    formula = y ~ poly(x, 24),
    alpha = 0.2,
    size = 0.5,
    show.legend = TRUE
  ) +
  facet_grid(. ~ Treatment) +
  scale_x_discrete(breaks = seq(0, 35, 5)) +
  scale_y_continuous(breaks = c(1, 1.2, 1.4, 1.6)) +
  labs(
    x = "Time (hours)", 
    y = "Fold confluency",
    title = "miR34a asRNA's role in growth regulation."
  ) + 
  guides(
    colour = guide_legend(title = "Cell line", override.aes = list(size = 3))
  )

markdown <- p +
  geom_label(
    data = filter(growth.lm, term == "`Cell line`lncTAM34a"),
    aes(x = 2, y = 1.64, label = paste0("estimate=", round(estimate, digits = 4))),
    fill = "white", label.size = 0, label.padding = unit(0.01, "lines"),
    position = position_dodge(width = 0.9), show.legend = FALSE,
    fontface = "bold", family = "Arial Unicode MS", size = 3, hjust = 0
  ) +
  geom_label(
    data = filter(growth.lm, term == "`Cell line`lncTAM34a"),
    aes(x = 2, y = 1.62, label = paste0("std.error=", round(std.error, digits = 4))),
    fill = "white", label.size = 0, label.padding = unit(0.01, "lines"),
    position = position_dodge(width = 0.9), show.legend = FALSE,
    fontface = "bold", family = "Arial Unicode MS", size = 3, hjust = 0
  ) +
  geom_label(
    data = filter(growth.lm, term == "`Cell line`lncTAM34a"),
    aes(x = 2, y = 1.6, label = paste0("p=", p.value)),
    fill = "white", label.size = 0, label.padding = unit(0.01, "lines"),
    position = position_dodge(width = 0.9), show.legend = FALSE,
    fontface = "bold", family = "Arial Unicode MS", size = 3, hjust = 0
  )

plotRmarkdown(markdown)

pdf <- p +
  geom_label(
    data = filter(growth.lm, term == "`Cell line`lncTAM34a"),
    aes(x = 2, y = 1.68, label = paste0("estimate=", round(estimate, digits = 4))),
    fill = "white", label.size = 0, label.padding = unit(0.01, "lines"),
    position = position_dodge(width = 0.9), show.legend = FALSE,
    fontface = "plain", family = "Arial Unicode MS", size = 2, hjust = 0
  ) +
  geom_label(
    data = filter(growth.lm, term == "`Cell line`lncTAM34a"),
    aes(x = 2, y = 1.64, label = paste0("std.error=", round(std.error, digits = 4))),
    fill = "white", label.size = 0, label.padding = unit(0.01, "lines"),
    position = position_dodge(width = 0.9), show.legend = FALSE,
    fontface = "plain", family = "Arial Unicode MS", size = 2, hjust = 0
  ) +
  geom_label(
    data = filter(growth.lm, term == "`Cell line`lncTAM34a"),
    aes(x = 2, y = 1.6, label = paste0("p=", p.value)),
    fill = "white", label.size = 0, label.padding = unit(0.01, "lines"),
    position = position_dodge(width = 0.9), show.legend = FALSE,
    fontface = "plain", family = "Arial Unicode MS", size = 2, hjust = 0
  )

figure <- plotPDF(pdf)

path <- file.path("~/GitHub/miR34a_asRNA_project/inst", fileMap(type = "pdf")["Figure 3c"][[1]])
ggsave(
  plot = figure,
  filename = path,
  device = cairo_pdf,
  height = 60,
  width = 83,
  units = "mm"
)
```

<div class="caption">
The effects of miR34a asRNA overexpression on growth in normal and starvation conditions in the PC3 prostate cancer cell line. Fold confluency (% confluency / Time 0 % confluency) is indicated on the y-axis. Colored lines indicate the polynomial regression model where mean fold confluency from 3 independant experiments is the dependant variable and time and cell line are independant variables. The grey shadows illustrate the 95% confidence interval for the regression. Estimates, standard error (std.error), and P values for the cell line covariate for each model are indicated in the upper left hand corner.
</div>

<br></br>

#### _Conclusions_

miR34a asRNA overexpression causes a minor decrease in cell growth under normal conditions (RPMI) although under conditions of cellular stress through starvation, this effect is increased.

<br></br><br></br>

```{r}
sessionInfo()
```